### 2021-12-02 manuscript.tex
Update image links and increment results [ci skip]
This build is based on
https://github.com/lubianat/quali_phd/commit/4244ca4800e49e86661876a45daa4ad73dc0ca6e.

This commit was created by the following CI build and job:
https://github.com/lubianat/quali_phd/commit/4244ca4800e49e86661876a45daa4ad73dc0ca6e/checks
https://github.com/lubianat/quali_phd/actions/runs/1530806454

This manuscript
(\href{https://lubianat.github.io/quali_phd/v/b00405796db47b9b3cdee9651d1d5273e977fd3e/}{permalink})
was automatically generated
from \href{https://github.com/lubianat/quali_phd/tree/b00405796db47b9b3cdee9651d1d5273e977fd3e}{lubianat/quali\_phd@b004057}
on December 2, 2021.

@@

\begin{itemize}
\tightlist
\item
  Wikidata as a knowledge graph for the life sciences
\end{itemize}

@@

In order to handle the literature reading necessary for this project, a framework was developed for reading. It consists of a set of Python scripts and a standard file structure. A file contains the list of articles to be read in markdown. Articles are represented as Wikidata QIDs which enables automatic information retrieval from Wikidata's structured systems. Articles were organized in 2 main sections, one for cell-type related articles and one for biocuration-related articles.

Besides the technical aspects of Wikidata Bib, the organized reading methodology included a discipline step of continued reading, with a target of 1-2 papers per section per day. Based on Umberto Eco's suggestion on How to Write a Thesis {[}\protect\hyperlink{ref-1HBVPtZGp}{121}{]} to develop a careful indexing system for literature, an index document was constructed containing the topics of interest for writing the thesis. The topics were added as plain text in the personalized note documents, allowing batch retrieval of articles of interest via the command line, using \texttt{grep} (https://en.wikipedia.org/w/index.php?title=Grep\&oldid=1039541979).

@@

Marker information was also recorded when explicitly mentioned and it will be added to Wikidata at a later step.

@@

As part of the research-action process, I have joined the Cell Ontology working group

\hypertarget{data-retrieval}{%
\subsection{Data retrieval}\label{data-retrieval}}

\begin{itemize}
\tightlist
\item
  SPARQL queries
\end{itemize}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

\begin{itemize}
\tightlist
\item
  Packages used in R and Python
\item
  For interacting with Wikidata
\end{itemize}

@@

\hypertarget{cell-disease-network-analysis}{%
\subsubsection{Cell-disease network analysis}\label{cell-disease-network-analysis}}

@@

In this period, we targeted the question: which cell type definition allows crafting coherent biological statements?
The goal was to not say what cell types \emph{are}, but what they can be for a consistent representation on an ontology or a knowledge graph, like Wikidata.
We avoided the dissection of the differences between persistent classes of cells (often called ``cell types'') or the transient, fugacious classes of cells (often called ``cell states'') (see ``Definition of cell identity'' section in {[}\protect\hyperlink{ref-pFijpXkl}{122}{]} for an example).
Even though such a distinction is an essential topic for theoretical research, it is not required to represent formally biomedical experiments.

To facilitate communication among life scientists, in a preprint derived from this PHD project{[}\protect\hyperlink{ref-uCE45Jh8}{123}{]}, we proposed, among other theoretical novelties, naming conventions for different cell types classes.
Much of the literature mixes cell types in one species (e.g., when dealing with a cell type as an evolutionary unit) or multiple species (e.g., in the Cell Ontology).
It is useful to distill these different concepts into names. Given the importance of the species' concept in biological classification {[}\protect\hyperlink{ref-AHGRV6KQ}{124}{]}, we derive a species-centric view on the naming of classes of cell types.
The four classes (Figure \ref{fig:classes}) we propose are as follows:

@@

Refining the different concepts around the notion of ``cell type'' is important, but will require decades before a reasonable consensus.
Here we adopt a liberal view of cell type, defining, for our purposes, a cell type as ``any class of cells described by a domain expert''.
By ``class'' we mean an abstract entity in the sense intended by the multilevel theory of conceptual modelling (referred as ``\emph{types}'' by Carvalho et al.~) {[}\protect\hyperlink{ref-eP47FILn}{125}{]}

@@

\begin{figure}
\hypertarget{fig:multilevel}{%
\centering
\includegraphics[width=1\textwidth,height=\textheight]{https://raw.githubusercontent.com/lubianat/multilevel_ontology_drawings/master/combination_human_cell.png}
\caption{Multileveltheory for cell types}\label{fig:multilevel}
}
\end{figure}

@@

PanglaoDB {[}\protect\hyperlink{ref-M0lqRYJb}{126}{]} {[}\protect\hyperlink{ref-T3PlZ3Vh}{127}{]} is a publically-available database that contains data and metadata on hundreds of single-cell RNA sequencing experiments.
It provides extensive information on cell types, genes, and tissues and cell type markers, obtained both via automatic and manual methods.
It also displays a rich web user interface for easy data acquisition, including database dumps for bulk downloads.

As of 17 June 2021, the article describing PanglaoDB had been cited 147 times.
Despite its use by the community, the database is on a 3-star category for Linked Open Data {[}\protect\hyperlink{ref-OGXYtGZ8}{128}{]} as it does not use the open semantic standards from W3C (RDF and SPARQL) needed for a 4-star rank, neither the links to external data via standard identifiers that make datasets 5-star.
Improving the data format toward W3C's gold standards is a valuable step in making biological knowledge FAIR (Findable, Accessible, Interoperable, and Reusable).
Thus, we aimed to provide a case study of making the core information of PandlaoDB available in a 5-star Linked Open Data Format while improving the modeling of the necessary concepts on Wikidata.

As of August 2020, Wikidata had 264 items being categorized as a ``cell type'', considerably less than in specialized cell catalogs, which count over two thousand cell types {[}\protect\hyperlink{ref-agnqfdk6}{45},\protect\hyperlink{ref-4AEy2xhQ}{129}{]}.
Strikingly, there were also 23 items categorized as ``instances of cell (Q7868)'' . This classification is imprecise, as an instance of cell would be an individual named cell from a single named individual.

@@

After obtaining approval from the owners of the database, we matched genes and cell types to Wikidata, and performed Wikidata queries to demonstrate the value of the approach. An overview of the process is shown in \ref{fig:iscb_intro}.
\includegraphics{images/poster_ISCB_intro}

@@

\begin{figure}
\hypertarget{fig:chat_marker}{%
\centering
\includegraphics{images/chat_marker.png}
\caption{Subset of the marker genes for item Q101405051 (human cholinergic neuron )}\label{fig:chat_marker}
}
\end{figure}

Now that we re-formatted the markers on PanglaoDB as Linked Open Data, we can make queries that were not possible before, including
federated queries with other biological databases, such as Uniprot {[}\protect\hyperlink{ref-HUJ9NzAb}{130}{]}
and Wikipathways {[}\protect\hyperlink{ref-AgOOCR24}{131}{]}.
Due to previous similar reconciliation projects, Wikidata already contains information about genes, including their relations to Gene Ontology (GO) terms.

@@

\begin{figure}
\hypertarget{fig:iscb_results}{%
\centering
\includegraphics{images/poster_ISCB_results.png}
\caption{SPARQL queries in Wikidata now harness information from Panglao DB. Queries with the above design were run on Wikidata. Results might change in real time with Wikidata updates by contributors A-C) Graphical representation of feasible SPARQL queries (https://w.wiki/yQ6, https://w.wiki/yQD and https://w.wiki/3HjX,). D) Sample result from query C.}\label{fig:iscb_results}
}
\end{figure}

@@

This work exemplifies the power of releasing Linked Open Data via Wikidata, and provides the biomedical community with the first semantically accessible, 5-star LOD dataset of cell markers, easily reachable from Wikidata's SPARQL Query Service (\url{https://query.wikidata.org/}).
It is not first case study of biomedical data integration to Wikidata (see {[}\protect\hyperlink{ref-mPoPwN77}{113}{]} for example.
Nevertheless, the differences among the articles in style and scope contribute to a richer ecosystem for possible contributor.
{]})
The work also paves the way for Wikidata reconciling of other databases for cell-type markers, such as CellMarker {[}\protect\hyperlink{ref-chGii6yw}{43}{]}, labome {[}\protect\hyperlink{ref-rhRRCtlA}{132}{]}, CellFinder {[}\protect\hyperlink{ref-4AEy2xhQ}{129}{]} and SHOGoiN/CELLPEDIA {[}\protect\hyperlink{ref-6uWWsiSq}{133}/{]}) (if proper authorization are given by the owners).
The approach we took here can in essence be applied to any knowledge set of public interest, providing a low-cost and low-barrier platform for sharing biocurated knowledge in gold standard format.

@@

@@

\begin{itemize}
\item
  Accountants have Double-entry bookkeeping (https://en.wikipedia.org/wiki/Double-entry\_bookkeeping), software developers have Test-driven development (https://en.wikipedia.org/wiki/Test-driven\_development).
\item
  Develop a professional way for coverage of large-scale revisions, inspired by Umberto Eco's How to Write a Thesis {[}\protect\hyperlink{ref-1HBVPtZGp}{121}{]} adapted to the digital environment, using version-control and semantic links.
\item
  Connect the reading framework with a Biocuration strategy to feed knowledge to Wikidata
\end{itemize}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-1HBVPtZGp}{}}%
\CSLLeftMargin{121. }
\CSLRightInline{\textbf{Come si fa una tesi di laurea} \url{https://www.wikidata.org/wiki/Q3684178}}

\leavevmode\vadjust pre{\hypertarget{ref-pFijpXkl}{}}%
\CSLLeftMargin{122. }
\CSLRightInline{\textbf{The Human Cell Atlas: Technical approaches and challenges.}
\CSLBlock{Chung Chau Hon, Jay W Shin, Piero Carninci, Michael JT Stubbington} \emph{Briefings in functional genomics} (2017-10-28) \url{https://www.wikidata.org/wiki/Q48563763}
\CSLBlock{DOI: \href{https://doi.org/10.1093/bfgp/elx029}{10.1093/bfgp/elx029}}}

\leavevmode\vadjust pre{\hypertarget{ref-uCE45Jh8}{}}%
\CSLLeftMargin{123. }
\CSLRightInline{\textbf{Towards a pragmatic definition of cell type}
\CSLBlock{Tiago Lubiana, Helder I Nakaya} \emph{Authorea, Inc.} (2021-01-04) \url{https://doi.org/ghrxwf}
\CSLBlock{DOI: \href{https://doi.org/10.22541/au.160979530.02627436/v1}{10.22541/au.160979530.02627436/v1}}}

\leavevmode\vadjust pre{\hypertarget{ref-AHGRV6KQ}{}}%
\CSLLeftMargin{124. }
\CSLRightInline{\textbf{PhyloCode} \url{https://www.wikidata.org/wiki/Q1189395}}

\leavevmode\vadjust pre{\hypertarget{ref-eP47FILn}{}}%
\CSLLeftMargin{125. }
\CSLRightInline{\textbf{Multi-level ontology-based conceptual modeling}
\CSLBlock{Victorio A Carvalho, João Paulo A Almeida, Claudenir M Fonseca, Giancarlo Guizzardi} \emph{Data and Knowledge Engineering} (2017-05-01) \url{https://www.wikidata.org/wiki/Q108926456}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.datak.2017.03.002}{10.1016/j.datak.2017.03.002}}}

\leavevmode\vadjust pre{\hypertarget{ref-M0lqRYJb}{}}%
\CSLLeftMargin{126. }
\CSLRightInline{\textbf{PanglaoDB - A Single Cell Sequencing Resource For Gene Expression Data} \url{https://panglaodb.se/index.html}}

\leavevmode\vadjust pre{\hypertarget{ref-T3PlZ3Vh}{}}%
\CSLLeftMargin{127. }
\CSLRightInline{\textbf{PanglaoDB: a web server for exploration of mouse and human single-cell RNA sequencing data}
\CSLBlock{Oscar Franzén, Li-Ming Gan, Johan LM Björkegren} \emph{Database} (2019) \url{https://doi.org/ggkzxr}
\CSLBlock{DOI: \href{https://doi.org/10.1093/database/baz046}{10.1093/database/baz046} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/30951143}{30951143} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6450036}{PMC6450036}}}

\leavevmode\vadjust pre{\hypertarget{ref-OGXYtGZ8}{}}%
\CSLLeftMargin{128. }
\CSLRightInline{\textbf{Linked Data - Design Issues} \url{https://www.w3.org/DesignIssues/LinkedData.html}}

\leavevmode\vadjust pre{\hypertarget{ref-4AEy2xhQ}{}}%
\CSLLeftMargin{129. }
\CSLRightInline{\textbf{CellFinder: a cell data repository}
\CSLBlock{Harald Stachelscheid, Stefanie Seltmann, Fritz Lekschas, Jean-Fred Fontaine, Nancy Mah, Mariana Lara Neves, Miguel A Andrade-Navarro, Ulf Leser, Andreas Kurtz} \emph{Nucleic Acids Research} (2013-12-03) \url{https://www.wikidata.org/wiki/Q28660708}
\CSLBlock{DOI: \href{https://doi.org/10.1093/nar/gkt1264}{10.1093/nar/gkt1264}}}

\leavevmode\vadjust pre{\hypertarget{ref-HUJ9NzAb}{}}%
\CSLLeftMargin{130. }
\CSLRightInline{\textbf{UniProt} \url{https://sparql.uniprot.org/sparql}}

\leavevmode\vadjust pre{\hypertarget{ref-AgOOCR24}{}}%
\CSLLeftMargin{131. }
\CSLRightInline{\textbf{Portal:Semantic Web - WikiPathways} \url{https://www.wikipathways.org/index.php/Portal:Semantic_Web}}

\leavevmode\vadjust pre{\hypertarget{ref-rhRRCtlA}{}}%
\CSLLeftMargin{132. }
\CSLRightInline{\textbf{Cell Markers}
\CSLBlock{Konstantin Yakimchuk} \emph{Materials and Methods} (2013-05-02) \url{https://doi.org/ghq494}
\CSLBlock{DOI: \href{https://doi.org/10.13070/mm.en.3.183}{10.13070/mm.en.3.183}}}

\leavevmode\vadjust pre{\hypertarget{ref-6uWWsiSq}{}}%
\CSLLeftMargin{133. }
\CSLRightInline{\textbf{SHOGoiN: Shogoin Human Omics database for the Generation of iPS and Normal cells} \url{https://stemcellinformatics.org/}}

>>>>

This manuscript
(\href{https://lubianat.github.io/quali_phd/v/4244ca4800e49e86661876a45daa4ad73dc0ca6e/}{permalink})
was automatically generated
from \href{https://github.com/lubianat/quali_phd/tree/4244ca4800e49e86661876a45daa4ad73dc0ca6e}{lubianat/quali\_phd@4244ca4}
on December 2, 2021.

@@

@@

To handle the literature reading necessary for this project, a framework was developed for reading and is described in details in the results section.
The framework is based on GitHub and includes Python scripts, a file organizing the reading list, and another documenting the reading history in RDF. Notes and additional information are saved in a GitHub repository, and the structured information powers a live website with analytics on the users recent readings. The source code for Wikidata Bib is available at \url{https://github.com/lubianat/wikidata_bib/tree/template} and notes on my readings can currently be accessed at \url{https://lubianat.github.io/wikidata_bib/}.

Additionaly, the methodology included a discipline of reading that entails the daily reading of 2 articles, one about ``cell types'' and another about ``biocuration''.
The articles are obtained by a mixed manual and automatic approach, including a la carte selection of articles to read alongside Wikidata queries for Cell, Nature, Science and eLife papers about single cell transcriptomics (query: \url{https://w.wiki/4LHr}) and for papers on biocuration (query: \url{https://w.wiki/4LHi}).
)

@@

@@

As part of the research-action process, I have joined the Cell Ontology working group.
I participate in the monthly meetings and sporadic workshops, learning and contributing to the discussions.
Additionaly, I contribute to the ontology development, actively engaging in the Cell Ontology GitHub repository (\url{https://github.com/obophenotype/cell-ontology}) and contributing with new terms and assertions.
I edit the ontology with the software for ontology editing Protégé v. 5.5.0 (\url{https://protege.stanford.edu/}).

@@

Status of cell type information on Wikidat was acessed via SPARQL queries combined with processing in python and is available at \url{https://colab.research.google.com/drive/1GvQXOs51_U8icdGMtKXMeLOXKM8pXWet\#scrollTo=szvBWI9zr_AA}

@@

In this period, we targeted the question: which cell type definition allows crafting coherent biological statements?
The goal was to not say what cell types \emph{are}, but what they can be for a consistent representation on an ontology or a knowledge graph, like Wikidata.
We avoided the dissection of the differences between persistent classes of cells (often called ``cell types'') or the transient, fugacious classes of cells (often called ``cell states'') (see ``Definition of cell identity'' section in {[}\protect\hyperlink{ref-pFijpXkl}{121}{]} for an example).
Even though such a distinction is an essential topic for theoretical research, it is not required to represent formally biomedical experiments.

To facilitate communication among life scientists, in a preprint derived from this PHD project{[}\protect\hyperlink{ref-uCE45Jh8}{122}{]}, we proposed, among other theoretical novelties, naming conventions for different cell types classes.
Much of the literature mixes cell types in one species (e.g., when dealing with a cell type as an evolutionary unit) or multiple species (e.g., in the Cell Ontology).
It is useful to distill these different concepts into names. Given the importance of the species' concept in biological classification {[}\protect\hyperlink{ref-AHGRV6KQ}{123}{]}, we derive a species-centric view on the naming of classes of cell types.
The four classes (Figure \ref{fig:classes}) we propose are as follows:

@@

Refining the different concepts around the notion of ``cell type'' is important, but will require decades before a reasonable consensus.
Here we adopt a liberal view of cell type, defining, for our purposes, a cell type as ``any class of cells described by a domain expert''.
By ``class'' we mean an abstract entity in the sense intended by the multilevel theory of conceptual modelling (referred as ``\emph{types}'' by Carvalho et al.~) {[}\protect\hyperlink{ref-eP47FILn}{124}{]}

@@

\begin{figure}
\hypertarget{fig:multilevel}{%
\centering
\includegraphics[width=0.85\textwidth,height=\textheight]{https://raw.githubusercontent.com/lubianat/multilevel_ontology_drawings/master/combination_human_cell.png}
\caption{Multileveltheory for cell types}\label{fig:multilevel}
}
\end{figure}

@@

PanglaoDB {[}\protect\hyperlink{ref-M0lqRYJb}{125}{]} {[}\protect\hyperlink{ref-T3PlZ3Vh}{126}{]} is a publically-available database that contains data and metadata on hundreds of single-cell RNA sequencing experiments.
It provides extensive information on cell types, genes, and tissues and cell type markers, obtained both via automatic and manual methods.
It also displays a rich web user interface for easy data acquisition, including database dumps for bulk downloads.

As of 17 June 2021, the article describing PanglaoDB had been cited 147 times.
Despite its use by the community, the database is on a 3-star category for Linked Open Data {[}\protect\hyperlink{ref-OGXYtGZ8}{127}{]} as it does not use the open semantic standards from W3C (RDF and SPARQL) needed for a 4-star rank, neither the links to external data via standard identifiers that make datasets 5-star.
Improving the data format toward W3C's gold standards is a valuable step in making biological knowledge FAIR (Findable, Accessible, Interoperable, and Reusable).
Thus, we aimed to provide a case study of making the core information of PandlaoDB available in a 5-star Linked Open Data Format while improving the modeling of the necessary concepts on Wikidata.

As of August 2020, Wikidata had 264 items being categorized as a ``cell type'', considerably less than in specialized cell catalogs, which count over two thousand cell types {[}\protect\hyperlink{ref-agnqfdk6}{45},\protect\hyperlink{ref-4AEy2xhQ}{128}{]}.
Strikingly, there were also 23 items categorized as ``instances of cell (Q7868)'' . This classification is imprecise, as an instance of cell would be an individual named cell from a single named individual.

@@

After obtaining approval from the owners of the database, we matched genes and cell types to Wikidata, and performed Wikidata queries to demonstrate the value of the approach. An overview of the process is shown in \ref{fig:iscb_intro}.
\includegraphics[width=0.85\textwidth,height=\textheight]{https://github.com/lubianat/fapesp_report_1/blob/main/content/images/poster_ISCB_intro.png?raw=true}

@@

\begin{figure}
\hypertarget{fig:chat_marker}{%
\centering
\includegraphics[width=0.85\textwidth,height=\textheight]{https://github.com/lubianat/fapesp_report_1/blob/main/content/images/chat_marker.png?raw=true}
\caption{Subset of the marker genes for item Q101405051 (human cholinergic neuron )}\label{fig:chat_marker}
}
\end{figure}

Now that we re-formatted the markers on PanglaoDB as Linked Open Data, we can make queries that were not possible before, including
federated queries with other biological databases, such as Uniprot {[}\protect\hyperlink{ref-HUJ9NzAb}{129}{]}
and Wikipathways {[}\protect\hyperlink{ref-AgOOCR24}{130}{]}.
Due to previous similar reconciliation projects, Wikidata already contains information about genes, including their relations to Gene Ontology (GO) terms.

@@

\begin{figure}
\hypertarget{fig:iscb_results}{%
\centering
\includegraphics[width=0.85\textwidth,height=\textheight]{https://github.com/lubianat/fapesp_report_1/blob/main/content/images/poster_ISCB_results.png?raw=true}
\caption{SPARQL queries in Wikidata now harness information from Panglao DB. Queries with the above design were run on Wikidata. Results might change in real time with Wikidata updates by contributors A-C) Graphical representation of feasible SPARQL queries (https://w.wiki/yQ6, https://w.wiki/yQD and https://w.wiki/3HjX,). D) Sample result from query C.}\label{fig:iscb_results}
}
\end{figure}

@@

This work exemplifies the power of releasing Linked Open Data via Wikidata, and provides the biomedical community with the first semantically accessible, 5-star LOD dataset of cell markers, easily reachable from Wikidata's SPARQL Query Service (\url{https://query.wikidata.org/}).
It is not first case study of biomedical data integration to Wikidata (see {[}\protect\hyperlink{ref-mPoPwN77}{113}{]} for example.
Nevertheless, the differences among the articles in style and scope contribute to a richer ecosystem for possible contributor.
{]})
The work also paves the way for Wikidata reconciling of other databases for cell-type markers, such as CellMarker {[}\protect\hyperlink{ref-chGii6yw}{43}{]}, labome {[}\protect\hyperlink{ref-rhRRCtlA}{131}{]}, CellFinder {[}\protect\hyperlink{ref-4AEy2xhQ}{128}{]} and SHOGoiN/CELLPEDIA {[}\protect\hyperlink{ref-6uWWsiSq}{132}/{]}) (if proper authorization are given by the owners).
The approach we took here can in essence be applied to any knowledge set of public interest, providing a low-cost and low-barrier platform for sharing biocurated knowledge in gold standard format.

@@

The contributions to cell types on Wikidata will be of most value if they are integrated to the current state-of-art of knowledge representation.
Arguably, the Cell Ontology is the current leading source of cell type identifiers in the context of the Human Cell Atlas project.{[}\protect\hyperlink{ref-qT8WxqjA}{48}{]}
Thus, it is crucial that data about cell types on Wikidata is connected to the Cell Ontology.

To start the improvement in the interplay of both databases, we proposed and got approval of a specific Wikidata identifier for the Cell Ontology, the ``Cell Ontology ID'' (\url{https://www.wikidata.org/wiki/Property:P7963}).
IDs can be added to Wikidata entities and connect them to external databases enabling integrative SPARQL queries.
Besides using the common Wikidata interface, one can crowd-curate identifiers via 3rd-party service, Mix'N'Match, which provides an user-friendly framework for connecting idenfier catalogs to Wikidata. {[}\protect\hyperlink{ref-JgiKEEdq}{133}/?p=114{]}, as seen in Figure \ref{fig:mixn_match_cl}.
Logically, we created a Mix'N'Match catalog for harmonizing Cell Ontology IDs to Wikidata (\url{https://mix-n-match.toolforge.org/\#/catalog/4719}), harnessing the community support for the task.

\begin{figure}
\hypertarget{fig:mixnmatch_cl}{%
\centering
\includegraphics[width=0.85\textwidth,height=\textheight]{https://pointstodots.files.wordpress.com/2021/09/image-17.png}
\caption{Mix'N'Match curation system}\label{fig:mixnmatch_cl}
}
\end{figure}

As of early December 2021, more than 700 Cell Ontology IDs have been manually matched to Wikidata.
The integration already enables queries that harness the previously existing information on Wikidata for Cell Ontology - based applications.
For example, one can query Wikidata items that have (1) a crossref to a CL ID (2) a picture in Wikimedia Commons (\url{https://w.wiki/4F6e}, Figure \ref{fig:cl_images}).
The different possibilities of mutual benefit between the Cell Ontology and Wikidata will continue to be explored in the next years of this PhD project.

\includegraphics{https://user-images.githubusercontent.com/7917951/137942026-7645f368-d62a-4434-be05-083555cf0757.png} \{\#fig:cl\_images width=``85\%''\}

@@

\begin{itemize}
\item
  Accountants have Double-entry bookkeeping (https://en.wikipedia.org/wiki/Double-entry\_bookkeeping), software developers have Test-driven development (https://en.wikipedia.org/wiki/Test-driven\_development).
\item
  Develop a professional way for coverage of large-scale revisions, inspired by Umberto Eco's How to Write a Thesis {[}\protect\hyperlink{ref-1HBVPtZGp}{134}{]} adapted to the digital environment, using version-control and semantic links.
\item
  Connect the reading framework with a Biocuration strategy to feed knowledge to Wikidata
\end{itemize}

@@

It consists of a set of Python scripts and a standard file structure. A file contains the list of articles to be read in markdown. Articles are represented as Wikidata QIDs which enables automatic information retrieval from Wikidata's structured systems. Articles were organized in 2 main sections, one for cell-type related articles and one for biocuration-related articles.

Besides the technical aspects of Wikidata Bib, the organized reading methodology included a discipline step of continued reading, with a target of 1-2 papers per section per day. Based on Umberto Eco's suggestion on How to Write a Thesis {[}\protect\hyperlink{ref-1HBVPtZGp}{134}{]} to develop a careful indexing system for literature, an index document was constructed containing the topics of interest for writing the thesis. The topics were added as plain text in the personalized note documents, allowing batch retrieval of articles of interest via the command line, using \texttt{grep} (https://en.wikipedia.org/w/index.php?title=Grep\&oldid=1039541979).

@@

\leavevmode\vadjust pre{\hypertarget{ref-pFijpXkl}{}}%
\CSLLeftMargin{121. }
\CSLRightInline{\textbf{The Human Cell Atlas: Technical approaches and challenges.}
\CSLBlock{Chung Chau Hon, Jay W Shin, Piero Carninci, Michael JT Stubbington} \emph{Briefings in functional genomics} (2017-10-28) \url{https://www.wikidata.org/wiki/Q48563763}
\CSLBlock{DOI: \href{https://doi.org/10.1093/bfgp/elx029}{10.1093/bfgp/elx029}}}

\leavevmode\vadjust pre{\hypertarget{ref-uCE45Jh8}{}}%
\CSLLeftMargin{122. }
\CSLRightInline{\textbf{Towards a pragmatic definition of cell type}
\CSLBlock{Tiago Lubiana, Helder I Nakaya} \emph{Authorea, Inc.} (2021-01-04) \url{https://doi.org/ghrxwf}
\CSLBlock{DOI: \href{https://doi.org/10.22541/au.160979530.02627436/v1}{10.22541/au.160979530.02627436/v1}}}

\leavevmode\vadjust pre{\hypertarget{ref-AHGRV6KQ}{}}%
\CSLLeftMargin{123. }
\CSLRightInline{\textbf{PhyloCode} \url{https://www.wikidata.org/wiki/Q1189395}}

\leavevmode\vadjust pre{\hypertarget{ref-eP47FILn}{}}%
\CSLLeftMargin{124. }
\CSLRightInline{\textbf{Multi-level ontology-based conceptual modeling}
\CSLBlock{Victorio A Carvalho, João Paulo A Almeida, Claudenir M Fonseca, Giancarlo Guizzardi} \emph{Data and Knowledge Engineering} (2017-05-01) \url{https://www.wikidata.org/wiki/Q108926456}
\CSLBlock{DOI: \href{https://doi.org/10.1016/j.datak.2017.03.002}{10.1016/j.datak.2017.03.002}}}

\leavevmode\vadjust pre{\hypertarget{ref-M0lqRYJb}{}}%
\CSLLeftMargin{125. }
\CSLRightInline{\textbf{PanglaoDB - A Single Cell Sequencing Resource For Gene Expression Data} \url{https://panglaodb.se/index.html}}

\leavevmode\vadjust pre{\hypertarget{ref-T3PlZ3Vh}{}}%
\CSLLeftMargin{126. }
\CSLRightInline{\textbf{PanglaoDB: a web server for exploration of mouse and human single-cell RNA sequencing data}
\CSLBlock{Oscar Franzén, Li-Ming Gan, Johan LM Björkegren} \emph{Database} (2019) \url{https://doi.org/ggkzxr}
\CSLBlock{DOI: \href{https://doi.org/10.1093/database/baz046}{10.1093/database/baz046} · PMID: \href{https://www.ncbi.nlm.nih.gov/pubmed/30951143}{30951143} · PMCID: \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6450036}{PMC6450036}}}

\leavevmode\vadjust pre{\hypertarget{ref-OGXYtGZ8}{}}%
\CSLLeftMargin{127. }
\CSLRightInline{\textbf{Linked Data - Design Issues} \url{https://www.w3.org/DesignIssues/LinkedData.html}}

\leavevmode\vadjust pre{\hypertarget{ref-4AEy2xhQ}{}}%
\CSLLeftMargin{128. }
\CSLRightInline{\textbf{CellFinder: a cell data repository}
\CSLBlock{Harald Stachelscheid, Stefanie Seltmann, Fritz Lekschas, Jean-Fred Fontaine, Nancy Mah, Mariana Lara Neves, Miguel A Andrade-Navarro, Ulf Leser, Andreas Kurtz} \emph{Nucleic Acids Research} (2013-12-03) \url{https://www.wikidata.org/wiki/Q28660708}
\CSLBlock{DOI: \href{https://doi.org/10.1093/nar/gkt1264}{10.1093/nar/gkt1264}}}

\leavevmode\vadjust pre{\hypertarget{ref-HUJ9NzAb}{}}%
\CSLLeftMargin{129. }
\CSLRightInline{\textbf{UniProt} \url{https://sparql.uniprot.org/sparql}}

\leavevmode\vadjust pre{\hypertarget{ref-AgOOCR24}{}}%
\CSLLeftMargin{130. }
\CSLRightInline{\textbf{Portal:Semantic Web - WikiPathways} \url{https://www.wikipathways.org/index.php/Portal:Semantic_Web}}

\leavevmode\vadjust pre{\hypertarget{ref-rhRRCtlA}{}}%
\CSLLeftMargin{131. }
\CSLRightInline{\textbf{Cell Markers}
\CSLBlock{Konstantin Yakimchuk} \emph{Materials and Methods} (2013-05-02) \url{https://doi.org/ghq494}
\CSLBlock{DOI: \href{https://doi.org/10.13070/mm.en.3.183}{10.13070/mm.en.3.183}}}

\leavevmode\vadjust pre{\hypertarget{ref-6uWWsiSq}{}}%
\CSLLeftMargin{132. }
\CSLRightInline{\textbf{SHOGoiN: Shogoin Human Omics database for the Generation of iPS and Normal cells} \url{https://stemcellinformatics.org/}}

\leavevmode\vadjust pre{\hypertarget{ref-JgiKEEdq}{}}%
\CSLLeftMargin{133. }
\CSLRightInline{\textbf{The Whelming › Tech, tools, and tribulations}
\CSLBlock{Scott Allan Wallick} \url{http://magnusmanske.de/wordpress/}}

\leavevmode\vadjust pre{\hypertarget{ref-1HBVPtZGp}{}}%
\CSLLeftMargin{134. }
\CSLRightInline{\textbf{Come si fa una tesi di laurea} \url{https://www.wikidata.org/wiki/Q3684178}}
